---
title: "Emergent Strategies for Strength Development"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---
```{r, include = F}
library(tidyverse)
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(ggridges)
library(scales)
library(stringr)
library(lubridate)
library(scales)
library(leaps)
library(TSA)
library(tseries)
library(astsa)
```







```{r,include=F}
trainingLog <- read.csv("data/trainingLog copy.csv")
trainingLog <- tibble::as_tibble(trainingLog)
trainingLog
```

```{r,include=F}
#1. make sure the dates are utilized as dates.
trainingLog$date <-mdy(trainingLog$date)
trainingLog

# the dates are now in ymd format and seen as a date, where blanks are NA

# fill the NA dates...
trainingLog <- trainingLog %>% fill(date)

#2. unite the columns that specify an exercise
#since i didn't use equipment 3, i'll get rid of it...
trainingLog$equipment3 <- NULL
      ##trainingLog <- trainingLog %>% mutate_if(is.factor, funs(na_if(.,""))) 
trainingLog <- trainingLog %>% unite("exercise",2:9, sep= " ")

#now I've got the exercises identified uniquely. Next, forward fill by setting all blank strings to NA, then using fill fn
trainingLog$exercise<- replace(trainingLog$exercise,which(trainingLog$exercise == "       "), NA)
trainingLog<- trainingLog %>% fill(exercise)
#now the data is forwardly filled, but there's annoying whitespace that needs trimming

## this will trim all the spaces with >1 space to just 1 space
trainingLog$exercise <- gsub("\\s+", " ", trainingLog$exercise) 

#this will trim all leading and trailing whitespace
trainingLog$exercise <- trimws(trainingLog$exercise) 

## now make the remaining values with blank strings NA so they're easier to deal with when plotting...
trainingLog <- trainingLog %>% mutate_if(is.factor, funs(na_if(.,""))) 

## replacing bogus RPE values (ones not conforming to the .5 interval system or <6)
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "9.4","9.5")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == ".",NA)
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "6.4","6.5")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "9.75","10")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "8.6","9.5")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "19","10")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "7.6","7.5")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "75","7.5")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "<5","<6")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "5","<6")
trainingLog$RPE<- replace(trainingLog$RPE,trainingLog$RPE == "5.5","<6")

#changing rpe to character vector
trainingLog$RPE<- as.character(trainingLog$RPE)



```

```{r,include=F}
# need to cluster the exercises into the following groups:
  #squating Movement, horizontal push, vertical push, deadlift movement, upper back, hamstring/glute isolation, tricep isolation
#squat movement - contains 'squat' or 'lunge'
squatIds <- c("squat","lunge")
#horiz press - contains 'bench press' | 'dip' | 'floor press' | 'spoto'
horizontalPressIds <- c("bench press", "dip", "floor press","spoto")
# vertical push - contains 'OHP'| 'ohp' | 'z-press' |'arnold press'
verticalPressIds <- c("ohp", "z-press", "arnold press")
#deadlift movement - contains 'DL'
DLIds <- c("dl","block pull", "rack pull")
# upper back - contains 'lat' | 'pulldown' |'row' | 'chinup' | 'pullup' 
upperBackIds <- c("lat","pulldown","row","chinup","pullup")
# hamstring/glute iso - contains 'ghr' | 'ham curl' | 'hip thrust' | 
hamstringIds<-c("ghr","ham curl")
gluteIds <- c("hip thrust")
# tricep iso - contains "OH tri" | "skullcrusher" 
tricepIds <- c("oh tri","skullcrusher")
ex <- str_to_lower(trainingLog$exercise)
trainingLog<- trainingLog %>% mutate(
  group = ifelse(str_detect(ex,horizontalPressIds[1]),"horizontalPress",
                ifelse(str_detect(ex,horizontalPressIds[2]),"horizontalPress",
                 ifelse(str_detect(ex,horizontalPressIds[3]),"horizontalPress",
                 ifelse(str_detect(ex,horizontalPressIds[4]),"horizontalPress",
                 ifelse(str_detect(ex,verticalPressIds[1]),"verticalPress",
                 ifelse( str_detect(ex,verticalPressIds[2]),"verticalPress",
                 ifelse(str_detect(ex,verticalPressIds[2]),"verticalPress",
                 ifelse(str_detect(ex,verticalPressIds[1]),"verticalPress",
                 ifelse(str_detect(ex,verticalPressIds[2]),"verticalPress",
                 ifelse(str_detect(ex,verticalPressIds[3]),"verticalPress",
                 ifelse(str_detect(ex,verticalPressIds[1]),"verticalPress",
                 ifelse(str_detect(ex,DLIds[1]),"DLMovement",
                 ifelse(str_detect(ex,DLIds[2]),"DLMovement",
                 ifelse(str_detect(ex,DLIds[3]),"DLMovement",
                 ifelse(str_detect(ex,upperBackIds[1]),"upperBack",
                 ifelse(str_detect(ex,upperBackIds[2]),"upperBack",
                 ifelse(str_detect(ex,upperBackIds[3]),"upperBack",
                 ifelse(str_detect(ex,upperBackIds[4]),"upperBack",
                 ifelse(str_detect(ex,upperBackIds[5]),"upperBack",
                 ifelse(str_detect(ex,hamstringIds[1]),"hamstringIsolation",
                 ifelse(str_detect(ex,hamstringIds[2]),"hamstringIsolation",
                 ifelse(str_detect(ex,gluteIds[1]),"gluteIsolation",
                 ifelse(str_detect(ex,squatIds[1]),"squatMovement",
                 ifelse(str_detect(ex,squatIds[2]),"squatMovement",
                 ifelse(str_detect(ex,tricepIds[1]),"tricepIsolation",
                 ifelse(str_detect(ex,tricepIds[2]),"tricepIsolation",NA))))))))))))))))))))))))))
)

```



```{r, warning=FALSE,include= F}
##estimated 1rm
rpeChart<- readr::read_csv("data/rpeChart copy.csv")
rpeChart<- rpeChart[1:7]
names(rpeChart)[1] <- "RPE"
rpeChart <- rpeChart %>% gather(2:7, key = "reps",value =  "percentages")
rpeChart$reps <- as.integer(rpeChart$reps)
rpeChart$RPE <- as.character(rpeChart$RPE)
# rpeChart <- rpeChart[-which(rpeChart$percentages<.82),]

e1rm<-c()
for(i in 1:nrow(trainingLog)){
  if(trainingLog$RPE[i] %in% rpeChart$RPE & trainingLog$reps[i] %in% rpeChart$reps 
    #why doesn't this condition work? & (subset(rpeChart,RPE==trainingLog$RPE[i] &reps==trainingLog$reps[i])[,"percentages",drop=T]) >.815 
    ){
    e1rm<-c(e1rm,
trainingLog$weight[i]/(subset(rpeChart,RPE==trainingLog$RPE[i] &reps==trainingLog$reps[i])[,"percentages",drop=T]))
  }
  else{
    e1rm<-c(e1rm,NA)
  }
}
trainingLog <- tibble::add_column(trainingLog,e1rm = e1rm)
rm(e1rm)
```

```{r,include=F}
#1.) relative stress index part 1 (impact of a set on successive set)
delta1rmPercent <- c()
for(i in 1:nrow(trainingLog)){
  nextI = i+1
  if (trainingLog$setNumber[nextI] ==1 | i == nrow(trainingLog)){delta1rmPercent <- c(delta1rmPercent,NA)
  }
  else{
    delta= ((trainingLog$e1rm[nextI] -trainingLog$e1rm[i])/trainingLog$e1rm[i])*100
    if(delta %>% is.na()){delta1rmPercent<-c(delta1rmPercent,NA)}
    else if(delta>0){ delta1rmPercent<-c(delta1rmPercent,0)}
    else{ delta1rmPercent<-c(delta1rmPercent,delta)}
  }
}

trainingLog <- (tibble::add_column(trainingLog,percentImpact1rm = delta1rmPercent))
```

```{r,include=F}
##RSI Part 2
stress<- trainingLog %>% filter(group %in% c("horizontalPress","verticalPress","squatMovement","DLMovement")) %>% group_by(group, reps, RPE) %>%  summarise(
  avgImpact = mean(percentImpact1rm, na.rm= T)
)
ggplot(stress, aes(x=as.numeric(RPE),y=-avgImpact)) +geom_point(aes(color = group))

```


```{r, include = F}
big3 <- trainingLog %>% filter(exercise %in% c("lowbar squat belt","bench press","conv DL", "conv DL belt"))
big3$exercise[big3$exercise == "conv DL"] <- "conv DL belt"

#big3p<- ggplot(big3,aes(x=date)) + geom_line(aes(y=e1rm,color = exercise)) + facet_wrap(~exercise, scales = "free_y",ncol=1) 
#plotly::ggplotly(big3p)
```

```{r, include =F}
big3Strict <- trainingLog %>% filter(exercise %in% c("lowbar squat belt","bench press","conv DL", "conv DL belt") & reps < 4)
big3Strict$exercise[big3Strict$exercise == "conv DL"] <- "conv DL belt"

#big3pStrict<- ggplot(big3Strict,aes(x=date)) + geom_line(aes(y=e1rm,color = exercise)) + facet_wrap(~exercise, scales = "free_y",ncol=1) 
#plotly::ggplotly(big3pStrict)
```

<br />

```{r,include=F}
#plots
#ridgeline plot showing distributions of weight for different of a few exercises
s1rm<-trainingLog %>% filter(exercise =="lowbar squat belt" & reps<6)
ggplot(s1rm,aes(x=date,y=e1rm)) + geom_line() + scale_x_date(date_breaks="2 months") + geom_smooth(method = "lm")

ggplot(s1rm,aes(x=e1rm,y=date,group = date)) + geom_density_ridges() + scale_x_continuous(name = "e1rm", limits = c(385,475),breaks = c(385,395,405,415,425,435,445,455,465,475,485,495)) +
  scale_y_date(breaks = date_breaks("2 months")) + 
  geom_vline(xintercept=mean(s1rm$e1rm,na.rm=T))


b1rm <- trainingLog %>% filter(exercise =="bench press" & reps<6)
ggplot(b1rm,aes(x=date,y=e1rm)) + geom_line() + scale_x_date(date_breaks="2 months") + geom_smooth(method = "lm")

ggplot(b1rm,aes(x=e1rm,y=date,group = date)) + geom_density_ridges() + scale_x_continuous(name = "e1rm", limits = c(295,365)) +
  scale_y_date(breaks = date_breaks("2 months")) + 
  geom_vline(xintercept=mean(b1rm$e1rm,na.rm=T))
#ridgeline plot of proj maxes for each set over time.
#ridgeline plot of absolute stress index of sets over time by exercise
#time series plots showing volume ovr time and proj 1rm over time
# time series plot showing avg weighgt over time by exercise
```

```{r, include =F}
big3WeeklyStats<- big3 %>% group_by(week=floor_date(date, "week"),exercise) %>% summarize(
  repsPerWeek = sum(reps),
  setsPerWeek = n(),
  avgIntensityByWeek = mean(weight/e1rm , na.rm=T)
) 
big3WeeklyLong <- gather(big3WeeklyStats,metric, value, -c(week,exercise))

  ggplot(big3WeeklyLong,aes(x=week,y=value)) + geom_line(aes(color=metric)) + facet_wrap(~metric,scales="free_y",ncol=1)
  
big3WeeklyStats<- big3WeeklyStats %>% mutate(
  repsPerWeekZ = (repsPerWeek-mean(repsPerWeek,na.rm=T))/sd(repsPerWeek,na.rm=T),
  setsPerWeekZ =(setsPerWeek-mean(setsPerWeek,na.rm=T))/sd(setsPerWeek,na.rm=T),
  avgIntZ = (avgIntensityByWeek-mean(avgIntensityByWeek,na.rm=T))/sd(avgIntensityByWeek,na.rm=T)
  )
big3Z<-big3WeeklyStats %>% dplyr::select(week,exercise,repsPerWeekZ,setsPerWeekZ,avgIntZ)
big3ZLong<- big3Z %>% gather(metric,value,-c(week,exercise,repsPerWeekZ))
ggplot(data=big3ZLong,aes(x=week,y=value)) + geom_line(aes(color=metric)) +facet_wrap(~exercise,ncol=1)  + scale_x_date(date_breaks="8 weeks")


```


```{r, include = F}
#devblock 1
devBlock1<- trainingLog %>% subset(date >= "2018-03-27" & date <= "2018-05-08")
devBlock1$exercise[devBlock1$exercise == "conv DL"] <- "conv DL belt"

weeklyE1rmblock1 <- devBlock1 %>% filter(reps ==1 & exercise %in% c("lowbar squat belt","bench press","conv DL belt")) %>% group_by(week=floor_date(date, "week")+2,exercise,group) %>%
  summarize(e1rm = max(e1rm)) 
benchAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% mean()
sqtAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% mean()
DLAvgb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% mean()
benchSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% sd()
sqtSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% sd()
DLSDb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% sd()

weeklyE1rmblock1<- weeklyE1rmblock1 %>% mutate(
    e1rmZ = ifelse(exercise == "bench press", (e1rm-benchAvgb1)/benchSDb1,
                     ifelse(exercise == "lowbar squat belt", (e1rm- sqtAvgb1)/sqtSDb1,
                            ifelse(exercise == "conv DL belt", (e1rm- DLAvgb1)/DLSDb1)))
  )


db1WeeklyStats <- devBlock1 %>% filter(group %in% c("horizontalPress","squatMovement","DLMovement")) %>% group_by(week=floor_date(date, "week")+2,group) %>% 
   summarize(sets=n(),
             reps = sum(reps),
             avgIntensity = mean(weight/e1rm,na.rm=T)
   )
db1WeeklyStats<- merge(db1WeeklyStats,weeklyE1rmblock1,by=c("week","group"))

variables <- c("avgIntensity","sdIntensity","avgSets","sdSets", "avgE1rm")
exercises <- c("bench press","lowbar squat belt", "conv DL belt")
group <- c("horizontalPress","squatMovement","DLMovement")
block<- c(replicate(length(variables),1))

avgAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% mean()
avgAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% mean()
avgAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% mean()
sdAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% sd()
sdAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% sd()
sdAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% sd()
##
avgSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% mean()
avgSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% mean()
avgSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% mean()
sdSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% sd()
sdSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% sd()
sdSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% sd()
#
db1WeeklyStats<- db1WeeklyStats %>% mutate(
  avgIntZ = ifelse(group =="horizontalPress",(avgIntensity-avgAvgIntPush)/sdAvgIntPush,
                   ifelse(group=="squatMovement",(avgIntensity-avgAvgIntSqt)/sdAvgIntSqt,
                   ifelse(group=="DLMovement",(avgIntensity-avgAvgIntPull)/sdAvgIntPull,NA))),
  setsZ = ifelse(group =="horizontalPress",(sets-avgSetsPush)/sdSetsPush,
                   ifelse(group=="squatMovement",(sets-avgSetsSqt)/sdSetsSqt,
                   ifelse(group=="DLMovement",(sets - avgSetsPull)/sdSetsPull,NA)))
   )
db1WeeklyStatsLong<-db1WeeklyStats %>% gather(metric,value, -c(week,group,exercise))

weeklyStatsDB1<- db1WeeklyStats
weeklyStatsDB1 <-cbind(weeklyStatsDB1, block = replicate(nrow(weeklyStatsDB1),1))
weeklyStatsDB1Long<- gather(weeklyStatsDB1,metric, value, -c(week,group,exercise,block))
```

```{r eval=FALSE, include=FALSE}
 
db1p1<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p2)
db1p1

db1p2<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p2)
db1p2

db1p3<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p3)
db1p3

db1p4<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","avgIntZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p4)
db1p4

```
```{r include=F}
#devblock2
devBlock2 <- trainingLog %>% subset(date >= "2018-05-21" & date <= "2018-08-10")
devBlock2$exercise[devBlock2$exercise == "conv DL"] <- "conv DL belt"

weeklyE1rmblock1 <- devBlock2 %>% filter(reps ==1 & exercise %in% c("lowbar squat belt","bench press","conv DL belt")) %>% group_by(week=floor_date(date, "week")+1,exercise,group) %>%
  summarize(e1rm = max(e1rm)) 
benchAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% mean()
sqtAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% mean()
DLAvgb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% mean()
benchSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% sd()
sqtSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% sd()
DLSDb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% sd()
initialSquat<- (weeklyE1rmblock1 %>% filter(week == "2018-05-21" & exercise=="lowbar squat belt"))$e1rm
initialBench<- (weeklyE1rmblock1 %>% filter(week == "2018-05-21" & exercise=="bench press"))$e1rm
initialDL <- (weeklyE1rmblock1 %>% filter(week == "2018-05-21" & exercise=="conv DL belt"))$e1rm
                
weeklyE1rmblock1<- weeklyE1rmblock1 %>% mutate(
    e1rmZ = ifelse(exercise == "bench press", (e1rm-benchAvgb1)/benchSDb1,
                     ifelse(exercise == "lowbar squat belt", (e1rm- sqtAvgb1)/sqtSDb1,
                            ifelse(exercise == "conv DL belt", (e1rm- DLAvgb1)/DLSDb1))),
    
    changeInE1rmFromBase = ifelse(exercise == "bench press", (e1rm-initialBench)/initialBench,
                     ifelse(exercise == "lowbar squat belt", (e1rm-initialSquat)/initialSquat,
                            ifelse(exercise == "conv DL belt", (e1rm- initialDL)/initialDL)))
  )


db1WeeklyStats <- devBlock2 %>% filter(group %in% c("horizontalPress","squatMovement","DLMovement")) %>% group_by(week=floor_date(date, "week")+1,group) %>% 
   summarize(sets=n(),
             reps = sum(reps),
             avgIntensity = mean(weight/e1rm,na.rm=T)
   )
db1WeeklyStats<- merge(db1WeeklyStats,weeklyE1rmblock1,by=c("week","group"))
avgAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% mean()
avgAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% mean()
avgAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% mean()
sdAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% sd()
sdAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% sd()
sdAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% sd()
##
avgSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% mean()
avgSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% mean()
avgSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% mean()
sdSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% sd()
sdSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% sd()
sdSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% sd()
#
db1WeeklyStats<- db1WeeklyStats %>% mutate(
  avgIntZ = ifelse(group =="horizontalPress",(avgIntensity-avgAvgIntPush)/sdAvgIntPush,
                   ifelse(group=="squatMovement",(avgIntensity-avgAvgIntSqt)/sdAvgIntSqt,
                   ifelse(group=="DLMovement",(avgIntensity-avgAvgIntPull)/sdAvgIntPull,NA))),
  setsZ = ifelse(group =="horizontalPress",(sets-avgSetsPush)/sdSetsPush,
                   ifelse(group=="squatMovement",(sets-avgSetsSqt)/sdSetsSqt,
                   ifelse(group=="DLMovement",(sets - avgSetsPull)/sdSetsPull,NA)))
   )
db1WeeklyStatsLong<-db1WeeklyStats %>% gather(metric,value, -c(week,group,exercise))

weeklyStatsDB2<- db1WeeklyStats
weeklyStatsDB2 <-cbind(weeklyStatsDB2, block = replicate(nrow(weeklyStatsDB2),2))
weeklyStatsDB2Long<- gather(weeklyStatsDB2,metric, value, -c(week,group,exercise,block))
```

```{r eval=FALSE, include=FALSE}
db1p5<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p5)
db1p5

db1p6<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p6)
db1p6

db1p7<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p7)
db1p7

db1p8<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","avgIntZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p8)
db1p8


db1p9<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("changeInE1rmFromBase","avgIntensity")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p9)
db1p5

```

```{r,include = F}
#devblock3
#devBlock3 %>% View()
devBlock3<- trainingLog %>% subset(date >= "2018-09-18" & date <= "2018-11-24")
#devBlock3 %>% filter(group=="squatMovement") %>% group_by(exercise,date,group) %>% summarize(sets=max(setNumber)) %>% view()
db3chart<- ggplot(devBlock3 %>% group_by(exercise,date,group) %>% summarize(sets=max(setNumber)),aes(y=sets,x=date))+geom_line(aes(color=group))
#plotly::ggplotly(db3chart)
db3chart

devBlock3$exercise[devBlock3$exercise == "conv DL"] <- "conv DL belt"

weeklyE1rmblock1 <- devBlock3 %>% filter(reps ==1 & exercise %in% c("lowbar squat belt","bench press","conv DL belt")) %>% group_by(week=floor_date(date, "week")+2,exercise,group) %>%
  summarize(e1rm = max(e1rm)) 
benchAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% mean()
sqtAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% mean()
DLAvgb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% mean()
benchSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% sd()
sqtSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% sd()
DLSDb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% sd()
initialSquat<- (weeklyE1rmblock1 %>% filter(week == "2018-09-18" & exercise=="lowbar squat belt"))$e1rm
initialBench<- (weeklyE1rmblock1 %>% filter(week == "2018-09-18" & exercise=="bench press"))$e1rm
initialDL <- (weeklyE1rmblock1 %>% filter(week == "2018-09-18" & exercise=="conv DL belt"))$e1rm
                
weeklyE1rmblock1<- weeklyE1rmblock1 %>% mutate(
    e1rmZ = ifelse(exercise == "bench press", (e1rm-benchAvgb1)/benchSDb1,
                     ifelse(exercise == "lowbar squat belt", (e1rm- sqtAvgb1)/sqtSDb1,
                            ifelse(exercise == "conv DL belt", (e1rm- DLAvgb1)/DLSDb1))),
    
    changeInE1rmFromBase = ifelse(exercise == "bench press", (e1rm-initialBench)/initialBench,
                     ifelse(exercise == "lowbar squat belt", (e1rm-initialSquat)/initialSquat,
                            ifelse(exercise == "conv DL belt", (e1rm- initialDL)/initialDL)))
  )


db1WeeklyStats <- devBlock3 %>% filter(group %in% c("horizontalPress","squatMovement","DLMovement")) %>% group_by(week=floor_date(date, "week")+2,group) %>% 
   summarize(sets=n(),
             reps = sum(reps),
             avgIntensity = mean(weight/e1rm,na.rm=T)
   )
db1WeeklyStats<- merge(db1WeeklyStats,weeklyE1rmblock1,by=c("week","group"))
avgAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% mean()
avgAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% mean()
avgAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% mean()
sdAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% sd()
sdAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% sd()
sdAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% sd()
##
avgSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% mean()
avgSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% mean()
avgSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% mean()
sdSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% sd()
sdSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% sd()
sdSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% sd()
#
db1WeeklyStats<- db1WeeklyStats %>% mutate(
  avgIntZ = ifelse(group =="horizontalPress",(avgIntensity-avgAvgIntPush)/sdAvgIntPush,
                   ifelse(group=="squatMovement",(avgIntensity-avgAvgIntSqt)/sdAvgIntSqt,
                   ifelse(group=="DLMovement",(avgIntensity-avgAvgIntPull)/sdAvgIntPull,NA))),
  setsZ = ifelse(group =="horizontalPress",(sets-avgSetsPush)/sdSetsPush,
                   ifelse(group=="squatMovement",(sets-avgSetsSqt)/sdSetsSqt,
                   ifelse(group=="DLMovement",(sets - avgSetsPull)/sdSetsPull,NA)))
   )
db1WeeklyStatsLong<-db1WeeklyStats %>% gather(metric,value, -c(week,group,exercise))

weeklyStatsDB3<- db1WeeklyStats
weeklyStatsDB3 <-cbind(weeklyStatsDB3, block = replicate(nrow(weeklyStatsDB3),3))
weeklyStatsDB3Long<- gather(weeklyStatsDB3,metric, value, -c(block,group,week,exercise))


```

```{r eval=FALSE, include=FALSE}
 
db1p10<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p10)
db1p10
db1p11<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p11)
db1p11
db1p12<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p12)
db1p12
db1p13<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","avgIntZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p13)
db1p13

db1p14<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("changeInE1rmFromBase","avgIntensity")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p14)
db1p14
```
```{r,include = F}
#devblock4
devBlock4 <- trainingLog %>% subset(date >= "2018-11-24" & date <= "2019-02-09")

devBlock4$exercise[devBlock4$exercise == "conv DL"] <- "conv DL belt"

weeklyE1rmblock1 <- devBlock4 %>% filter(reps ==1 & exercise %in% c("lowbar squat belt","bench press","conv DL belt")) %>% group_by(week=floor_date(date,"week",week_start =6),exercise,group) %>%
  summarize(e1rm = max(e1rm)) 
benchAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% mean()
sqtAvgb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% mean()
DLAvgb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% mean()
benchSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "bench press"))$e1rm %>% sd()
sqtSDb1 <- (weeklyE1rmblock1 %>% filter(exercise == "lowbar squat belt"))$e1rm %>% sd()
DLSDb1<- (weeklyE1rmblock1 %>% filter(exercise == "conv DL belt"))$e1rm %>% sd()
initialSquat<- (weeklyE1rmblock1 %>% filter(week == "2018-11-24" & exercise=="lowbar squat belt"))$e1rm
initialBench<- (weeklyE1rmblock1 %>% filter(week == "2018-11-24" & exercise=="bench press"))$e1rm
initialDL <- (weeklyE1rmblock1 %>% filter(week == "2018-11-24" & exercise=="conv DL belt"))$e1rm
                
weeklyE1rmblock1<- weeklyE1rmblock1 %>% mutate(
    e1rmZ = ifelse(exercise == "bench press", (e1rm-benchAvgb1)/benchSDb1,
                     ifelse(exercise == "lowbar squat belt", (e1rm- sqtAvgb1)/sqtSDb1,
                            ifelse(exercise == "conv DL belt", (e1rm- DLAvgb1)/DLSDb1))),
    
    changeInE1rmFromBase = ifelse(exercise == "bench press", (e1rm-initialBench)/initialBench,
                     ifelse(exercise == "lowbar squat belt", (e1rm-initialSquat)/initialSquat,
                            ifelse(exercise == "conv DL belt", (e1rm- initialDL)/initialDL)))
  )


db1WeeklyStats <- devBlock4 %>% filter(group %in% c("horizontalPress","squatMovement","DLMovement")) %>% group_by(group,week=floor_date(date, "week",week_start = 6)) %>% 
   summarize(sets=n(),
             reps = sum(reps),
             avgIntensity = mean(weight/e1rm,na.rm=T)
   )
db1WeeklyStats<- merge(db1WeeklyStats,weeklyE1rmblock1,by=c("week","group"))
avgAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% mean()
avgAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% mean()
avgAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% mean()
sdAvgIntPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$avgIntensity %>% sd()
sdAvgIntPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$avgIntensity %>% sd()
sdAvgIntSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$avgIntensity %>% sd()
##
avgSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% mean()
avgSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% mean()
avgSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% mean()
sdSetsPush <- (db1WeeklyStats %>% filter(group == "horizontalPress"))$sets %>% sd()
sdSetsPull <- (db1WeeklyStats %>% filter(group == "DLMovement"))$sets %>% sd()
sdSetsSqt <- (db1WeeklyStats %>% filter(group == "squatMovement"))$sets %>% sd()
#
db1WeeklyStats<- db1WeeklyStats %>% mutate(
  avgIntZ = ifelse(group =="horizontalPress",(avgIntensity-avgAvgIntPush)/sdAvgIntPush,
                   ifelse(group=="squatMovement",(avgIntensity-avgAvgIntSqt)/sdAvgIntSqt,
                   ifelse(group=="DLMovement",(avgIntensity-avgAvgIntPull)/sdAvgIntPull,NA))),
  setsZ = ifelse(group =="horizontalPress",(sets-avgSetsPush)/sdSetsPush,
                   ifelse(group=="squatMovement",(sets-avgSetsSqt)/sdSetsSqt,
                   ifelse(group=="DLMovement",(sets - avgSetsPull)/sdSetsPull,NA)))
   )
db1WeeklyStatsLong<-db1WeeklyStats %>% gather(metric,value, -c(week,group,exercise))

weeklyStatsDB4<- db1WeeklyStats
weeklyStatsDB4 <-cbind(weeklyStatsDB4, block = replicate(nrow(weeklyStatsDB4),4))
weeklyStatsDB4Long<- gather(weeklyStatsDB4,metric, value, -c(week,group,exercise,block))


```

```{r eval=FALSE, include=FALSE}
 
db1p15<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p15)
db1p15
db1p16<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("avgIntZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p16)
db1p16
db1p17<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","e1rmZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p17)
db1p17
db1p18<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("setsZ","avgIntZ")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p18)
db1p18


db1p19<- ggplot(db1WeeklyStatsLong %>% filter(metric %in%c("changeInE1rmFromBase","avgIntensity")),aes(x=week,y=value)) + geom_line(aes(group=metric,color=metric))+facet_wrap(~exercise,ncol=1) 
#plotly::ggplotly(db1p19)
db1p19
```


# Using data at the meso-cycle/block level.

This is where the fun begins. A few months of training data and a couple valid assumptions can contain multiple insights to aid in future training programming for an individual. Here is an example block analysis for my own training data. I'll focus on the three main movements to keep it simple — the squat, the bench press, and the deadlift. 

```{r,echo=F}
d1<- ggplot(weeklyStatsDB1,aes(x=week,y=e1rm)) + geom_line(aes(color=exercise))+geom_point(aes(color=exercise))+facet_wrap(~exercise,scales="free_y")
plotly::ggplotly(d1) 
#d1
#sets per week facet by exercise color by block
 d1 <- ggplot(weeklyStatsDB1,aes(x=week,y=sets)) + geom_line(aes(color=exercise))+geom_point(aes(color=exercise)) + facet_wrap(~exercise,scales="free")
 #plotly::ggplotly(d1)
 d1
 d1<- ggplot(weeklyStatsDB1Long %>% filter(metric %in% c("e1rmZ","avgIntZ","setsZ")),aes(x=week,y=value)) + geom_line(aes(color = metric)) + geom_point(aes(color=metric)) +facet_wrap(~exercise,scales = "free_y") + ggtitle("Deadlift")
#plotly::ggplotly(d1)
d1
```

Potential insights: 

* 1.) natural "peak" occurs approximately 4-5 weeks into the block across these lifts. 

* 2.) The stall/decrease may have occurred for one of two reasons (most likely)
    
    * too much volume --> need to drop volume for a week or two and recover. (I tried this, and it didn't work)
    
    * too little volume --> not enough of a stimulus, increase volume. (Since option 1 didn't work, I tried this in the subsequent block)
            
Leading into the next block, I decided to increase the sets/week and decrease intensity accordingly. In the 3rd block, I tried to keep most variables similar to the second, except I "front loaded" volume on the squat (performed most of the high volume weeks in the first portion of the block). I also increased the bench volume. Let's see the results of adjusting the training parameters in a meta-block review. I'll examine the only the squat for the sake of brevity...


# Using data at the meta-block level

The greatest variables to check are volume and intensity, followed by frequency (provided exercise selection is relatively specific). We look at the following metrics

* e1rm over time for each exercise
* weekly volume over time in terms of sets and reps by exercise grouping (only including the squat variants to keep the example minimal)
* average weekly intensity as a proportion of the estimated one rep max (only including the squat variants to keep the example minimal)

```{r,include=F}
weeklyStatsAllBlocks <-weeklyStatsDB1 %>%  plyr::rbind.fill(weeklyStatsDB2) %>% plyr::rbind.fill(weeklyStatsDB3) %>% plyr::rbind.fill(weeklyStatsDB4)
weeklyStatsAllBlocksLong <-gather(weeklyStatsAllBlocks, metric, value,-c(week,group,exercise,block))
```

```{r,include=F,warning=F,message=F}
devBlockAll<-devBlock1 %>% mutate(block=1) %>%  plyr::rbind.fill(devBlock2 %>% mutate(block=2)) %>% plyr::rbind.fill(devBlock3 %>% mutate(block=3)) %>% plyr::rbind.fill(devBlock4 %>% mutate(block=4)) %>% group_by(block) %>%  mutate(t=date-min(date))
devBlockAll<- devBlockAll %>% group_by(block) %>% mutate(week=as.Date(floor_date(date,"week",week_start=wday(min(date),week_start=1)))) %>% mutate( tWeek=difftime(week,min(week),units="weeks"))
weekLevel<-devBlockAll %>% filter(group %in% c("DLMovement","horizontalPress","squatMovement"))  %>% group_by(group,block,week,tWeek) %>% summarize(numSets=n(),numReps=sum(reps),avgWeeklyIntensity=sum(reps * weight/mean(e1rm,na.rm=T) )/numReps, sdWeeklyIntensity=sqrt( sum(reps*(weight/mean(e1rm,na.rm=T) - avgWeeklyIntensity)^2/(numReps-1))) )
#weekLevel
#devBlockAll %>% group_by(block,week,group) %>% mutate(totalIntensity=reps*(weight/mean(e1rm))) %>% summarize(avgInt=totalIntensity/sum(reps))
weekLevelLong<-weekLevel %>% filter(group=="squatMovement") %>% gather(key="metric",value="val",c(numSets,numReps,avgWeeklyIntensity))

weekAvgs<-weekLevel %>% group_by(group,block) %>% summarize(avgSets=mean(numSets,na.rm=T),sdSets=sd(numSets,na.rm = T),meanReps=mean(numReps,na.rm=T),avgInt=mean(avgWeeklyIntensity,na.rm=T),sdInt=sd(avgWeeklyIntensity,na.rm=T),sdReps=sd(numReps,na.rm = T))
  
weeklyStatsAllBlocks <- weeklyStatsAllBlocks %>% group_by(block) %>% mutate(t=week-min(week))
sessLevel<-devBlockAll %>% filter(group %in% c("DLMovement","horizontalPress","squatMovement")) %>% group_by(group,block,t) %>% summarize(numSets=n(),numReps=sum(reps),avgSessionIntensity=mean(weight,na.rm=T)/mean(e1rm,na.rm=T))
```

```{r,include=F}
sessLevelLong<- sessLevel %>% gather(key="metric",value = "val",c(numSets,numReps,avgSessionIntensity))
tsp<- ggplot(sessLevelLong %>% filter(group=="squatMovement"),aes(y=val,x=t,color=as.factor(block)))+geom_line()+geom_point()+facet_wrap(~metric,scales="free")
#plotly::ggplotly(tsp)
tsp
sessionAvgs<- devBlockAll %>% filter(group %in% c("DLMovement","horizontalPress","squatMovement")) %>% group_by(group,block) %>% summarize(numSets=n(),numReps=sum(reps),numSessions=length(unique(date))) %>% mutate(avgRPS=numReps/numSessions,avgSPS=numSets/numSessions)


```

```{r,echo=F}
#delta e1rm facet by exercise color by block
pe1rm2<-ggplot(weeklyStatsAllBlocks,aes(x=week,y=e1rm)) + geom_line(aes(color=as.factor(block)))+geom_point(aes(color=as.factor(block))) + facet_wrap(~exercise,scales = "free",ncol=2) + scale_y_continuous(limits = c(300,550))
#pe1rm2
plotly::ggplotly(pe1rm2) 
```

```{r,echo=F,message=F,warning=F}

# plotly::ggplotly(
#   ggplot(weekLevel %>% filter(group=="squatMovement"),aes(y=avgWeeklyIntensity,x=tWeek,color=as.factor(block)))+geom_line()+geom_point() + geom_hline(data=weekAvgs,yintercept=avgInt)
# )

ggplot(weekLevel %>% filter(group=="squatMovement"),aes(y=avgWeeklyIntensity,x=tWeek,color=as.factor(block)))+geom_line()+geom_point() + geom_hline(data = weekAvgs %>% filter(group=="squatMovement"),aes( yintercept=avgInt,color=as.factor(block)),linetype="dashed")

#+geom_errorbar(aes(ymin=avgWeeklyIntensity-sdWeeklyIntensity, ymax=avgWeeklyIntensity+sdWeeklyIntensity), width=.1) 
    

ggplot(weekLevel %>% filter(group=="squatMovement"),aes(y=numSets,x=tWeek,color=as.factor(block)))+geom_line()+geom_point() +geom_hline(data = weekAvgs %>% filter(group=="squatMovement"),aes( yintercept=avgSets,color=as.factor(block)),linetype="dashed")
    


ggplot(weekLevel %>% filter(group=="squatMovement"),aes(y=numReps,x=tWeek,color=as.factor(block)))+geom_line()+geom_point() + geom_hline(data = weekAvgs %>% filter(group=="squatMovement"),aes( yintercept=meanReps,color=as.factor(block)),linetype="dashed")# + geom_hline(data = weekAvgs %>% filter(group=="squatMovement"),aes(yintercept=meanReps+sdReps,color=as.factor(block)),linetype="dotted") #+ geom_hline(data = weekAvgs %>% filter(group=="squatMovement"),aes(yintercept=meanReps-sdReps,color=as.factor(block)),linetype="dotted")


```

insights from comparing blocks 1 to 2 and 3...

 * Block 1 had insufficient volume and there was no need to keep intensity so high.
 
Insights from comparing blocks 2 and 3...

 * Both blocks were very successful (both projected e1rms increase by ~30 lbs).
 * Block 3 was successful AND predictable in that I hit the highest e1rm of the program on the exact day I planned to do so. The "peak" was precise. 
 * Sets per week were fairly similar (just shift block 3 two weeks to the right and they line up almost perfectly... this was by design)
 * intensity stayed above its average for most weeks in block 2, while block 3 stayed below its average for most weeks.
 * I managed to successfully slightly increase volume and intensity over the whole of the block while recovering adequately in block 3.
 
Going forward...

 * Be patient, wait until the last few weeks and then spike intensity sharply.
 * It works well when volume & reps remain higher "most the time", and intensity remains lower until it's time for the final peak. This might be a hint to stay at lower intensity ranges for a longer period of time in the future, or attempt to drop the intensity even further and have it climb from 72 percent unstead of from 77 percent.
 * It's okay to keep squatting into the "max-out" week. I actually managed to perform better on the final week with a less severe drop in volume.
 * 1 experiment: try to keep the same avg intensity,reps, and sets per week, but instead of front-loading volume (high week to week variance, low within the week variance), try to increase the within week variance and decrease week to week variance.
 
 